---
title: "RProject-2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE,warning = F)
```

install.packages("sqldf")
install.packages("rgdal")
install.packages("stringr")
install.packages("rgeos")
install.packages("dplyr")
install.packages("Hmisc")


```{r Set Working Directory}
setwd("F:/MSCSNyuPoly/Fall2016/FoundationsOfDataScience/Project/Work")
```

```{r Read Data}
#Read data
data2015 <- read.table("https://s3-us-west-2.amazonaws.com/fds2016sahil/data.txt")
#Total records in dataset
nrow(data2015)
data2015$Hearing.Status <- as.character(data2015$Hearing.Status)
data2015$Hearing.Result <- as.character(data2015$Hearing.Result)
data2015$Decision.Date <- as.Date(data2015$Decision.Date, format = "%m/%d/%Y")
data2015 <- subset(data2015, Total.Violation.Amount != 0)

```

```{r descriptive statistics of invalid zip code}
#Check number of invalid zip codes present
zipCodeDf <- data.frame(as.character(data2015$Violation.Location..Zip.Code.))
colnames(zipCodeDf) <- "zipCode"

zipCodeLength <- do.call(data.frame, aggregate(cbind(count = zipCode) ~ nchar(as.character(zipCodeDf$zipCode)), data = zipCodeDf, FUN = function(x) c({NROW(x)})))
colnames(zipCodeLength) <- c("ZipCodeLength", "count")
zipCodeLength
```



```{r optimum amount complete data}
#Select only Total Violation Amount column
data2015Amt <- subset(data2015, Total.Violation.Amount != 0, select = c("Total.Violation.Amount"))
#Total number of violations
ttlViolation <- nrow(data2015Amt)
#Aggregation to find percentage occurenece of each penalty amount
options(scipen = 999)  #To disable scietific notation
ttlAgg <- do.call(data.frame, aggregate(cbind(count = Total.Violation.Amount) ~ Total.Violation.Amount, data = data2015Amt, FUN = function(x) c({round((NROW(x) / ttlViolation) * 100, digits = 4)}, {NROW(x)})))
colnames(ttlAgg) <- c("Total.Violation.Amount", "ttlPercent", "ttlCnt")
```



```{r optimum amount paid data}
#Select data where Hearing Status is PAID IN FULL and Decision Date is blank. It means that amount has been paid in full before taking any decision
data2015Paid <- subset(data2015, Hearing.Status == "PAID IN FULL" & is.na(Decision.Date) & Total.Violation.Amount != 0, select = c("Total.Violation.Amount"))
#Aggregation  to find percentage occurence of penalty where user has paid fully
paidAgg <- do.call(data.frame, aggregate(cbind(count = Total.Violation.Amount) ~ Total.Violation.Amount, data = data2015Paid, FUN = function(x)c({round((NROW(x) / ttlViolation) * 100, digits = 4)}, {NROW(x)})))
colnames(paidAgg) <- c("Total.Violation.Amount", "paidPercentFromTotal", "paidCount")
```


```{r join optimum amount}
#Left join of ttl Percentage and paid percentage
mergeData <- merge(x = ttlAgg, y = paidAgg, by = "Total.Violation.Amount", all.x = TRUE)
#Replace NA with 0
mergeData <- data.frame(apply(mergeData, 2, function(x){replace(x, is.na(x), 0)}))
#Calculate percentage violations per amount where penalty was paid before decision
mergeData$paidPercentPerPenalty <- (mergeData[,"paidCount"] / mergeData[,"ttlCnt"]) * 100
head(mergeData)

```

```{r plot optimum amount}

#Barplot
rownames(mergeData) <- mergeData$Total.Violation.Amount
barplot(t(as.matrix(mergeData[,c(3, 5)])), col=c("darkblue","red"), legend=c("Total", "Paid"),ylim=c(0, 320000), xlab = "Amount", main = "Amount vs Count of Total & Amount Paid before Decision", beside=TRUE)
```
We are not able to see proper result because of large data. Even with barplot we are not able to visualize properly. We can divide data into bins to get overall estimation.

```{r divide into bins}
library(Hmisc)
mergeData$amountGrp <- cut2(mergeData$Total.Violation.Amount, c(101, 1001, 2501, 5001, 10001, 100000 ))
mergeDataAgg <- aggregate(mergeData[,c("ttlCnt","paidCount")], by=list(mergeData$amountGrp), "sum")
colnames(mergeDataAgg) <- c("Violation.Amount.Grp",  "ttlCnt", "paidCount")

barPlotData <- t(as.matrix(mergeDataAgg[,c(2,3)]))
names(barPlotData) <- mergeDataAgg[,c(1)]
bar <- barplot(barPlotData, col=c("darkblue","red"), legend=c("Total", "Paid"),ylim=c(0, 420642),ylab = "Count", xlab = "", main = "Amount vs Count of Total & Amount Paid before Decision", beside=TRUE)
axis(1, at=bar[1,], labels=mergeDataAgg[,c(1)], font = 2, cex.axis=0.7, las=3)
```


We can see from plot that maximum number of times amount was paid before decision when penalty amount was less than or equal to 100.

So lets zoom in there.
```{r zoom in}
barplot(t(as.matrix(mergeData[mergeData$Total.Violation.Amount <= 100,c(3, 5)])), col=c("darkblue","red"),ylim=c(0, 320000), xlab = "Amount", main = "Amount vs Count of Total & Amount Paid before Decision", beside=TRUE)
```


We can see that for lower penalty amount there is more percentage to pay penalty beofre decison. It is highest for penalty amount $25.



```{r Revenue Lost}
#How much revenue is lost on defaulted penalties
copyData2015 <- data2015
#Extract month from violation date
copyData2015$Violation.Date <- format(as.Date(as.character(copyData2015$Violation.Date), "%m/%d/%Y"), "%m")
#Select defaulted violations
data15MonthDefaulted <- subset(copyData2015, (Hearing.Status == "DOCKETED" | Hearing.Status == "DEFAULTED") & Hearing.Result == "DEFAULTED", select = c("Violation.Date", "Total.Violation.Amount"))

#Select subset of data for all violations
data15Month <- subset(copyData2015, select = c("Violation.Date", "Total.Violation.Amount"))

#Aggregate default violations
dataMonthDeafaultedAgg <- do.call(data.frame, aggregate(cbind(sum = Total.Violation.Amount) ~ Violation.Date, data = data15MonthDefaulted, FUN = sum))
colnames(dataMonthDeafaultedAgg) <- c("Violation.Month", "DefaultedSum")
#Aggregate all violations
data15MonthAgg <- do.call(data.frame, aggregate(cbind(sum = Total.Violation.Amount) ~ Violation.Date, data = data15Month, FUN = sum))
colnames(data15MonthAgg) <- c("Violation.Month", "ttlSum")
#Join and calculate percentage
mergeData <- merge(x = data15MonthAgg, y = dataMonthDeafaultedAgg, by = "Violation.Month", all.x = TRUE)
mergeData$Violation.Month <- as.character(mergeData$Violation.Month)
(mergeData)

#Plot bargraph
rownames(mergeData) <- mergeData$Violation.Month
barplot(t(as.matrix(mergeData[,c(2,3)])), xlab = "Month", col=c("darkblue","red"), main = "Monthwise Total Revenue and lost revenue", beside=TRUE)

legend("topleft", legend=c("Total", "Defaulted"), pch=20, col=c("darkblue","red"), horiz=TRUE, bty='n', cex=0.8)
```





# Find percentage of defaulted violations per agency, violationLoc, Penalty, Respondent Location
```{r aggregate on diff features}
copyData2015 <- data2015
#Select defaulted violations
data15Defaulted <- subset(copyData2015, (Hearing.Status == "DOCKETED" | Hearing.Status == "DEFAULTED") & Hearing.Result == "DEFAULTED", select = c("Charge..1..Code.Description", "Violation.Location..Zip.Code.", "Respondent.Address..Zip.Code.", "Total.Violation.Amount"))
#Select subset of data for all violations
data15All <- subset(copyData2015, select = c("Charge..1..Code.Description", "Violation.Location..Zip.Code.", "Respondent.Address..Zip.Code.", "Total.Violation.Amount"))

#Aggregate default violations
dataDeafaultedAgg <- do.call(data.frame, aggregate(cbind(DefaultedCount = Total.Violation.Amount) ~ (Charge..1..Code.Description + Violation.Location..Zip.Code. + Respondent.Address..Zip.Code. + Total.Violation.Amount), data = data15Defaulted, FUN = function(x)c({NROW(x)})))

#Aggregate all violations
data15AllAgg <- do.call(data.frame, aggregate(cbind(TotalCount = Total.Violation.Amount) ~ (Charge..1..Code.Description + Violation.Location..Zip.Code. + Respondent.Address..Zip.Code. + Total.Violation.Amount), data = data15All, FUN = function(x)c({NROW(x)})))


#Group by all columns
mergeData <- merge(x = data15AllAgg, y = dataDeafaultedAgg, by = c("Charge..1..Code.Description", "Violation.Location..Zip.Code.", "Respondent.Address..Zip.Code.", "Total.Violation.Amount"), all.x = TRUE)
mergeData <- data.frame(apply(mergeData, 2, function(x){replace(x, is.na(x), 0)}))
mergeData$TotalCount <- as.numeric(as.character(mergeData$TotalCount))
mergeData$DefaultedCount <- as.numeric(as.character(mergeData$DefaultedCount))
mergeData$Violation.Location..Zip.Code. <- (as.character(mergeData$Violation.Location..Zip.Code.))
mergeData$Respondent.Address..Zip.Code. <- (as.character(mergeData$Respondent.Address..Zip.Code.))
```

```{r agg by violation zip code}
#Agg by Violation zip code
mergeDataAgg <- aggregate(mergeData[,c("TotalCount", "DefaultedCount")], by = list(mergeData$Violation.Location..Zip.Code.), "sum")
colnames(mergeDataAgg) <-c("Violation.Location..Zip.Code.", "TotalCount", "DefaultedCount")
mergeDataAgg <- mergeDataAgg[(nchar(mergeDataAgg$Violation.Location..Zip.Code.) == 5 & mergeDataAgg$TotalCount >= 100),]
mergeDataAgg$Percentage <- (mergeDataAgg$DefaultedCount/ mergeDataAgg$TotalCount) * 100
top10 <- head(mergeDataAgg[order(mergeDataAgg$Percentage, decreasing = T), c(1,4)], 10)
top10
```

```{r aggregate by respondent zip code}
#Agg by respondent Zip code
mergeDataAgg <- aggregate(mergeData[,c("TotalCount", "DefaultedCount")], by = list(mergeData$Respondent.Address..Zip.Code.), "sum")
colnames(mergeDataAgg) <-c("Respondent.Address..Zip.Code.", "TotalCount", "DefaultedCount")
mergeDataAgg <- mergeDataAgg[(nchar(mergeDataAgg$Respondent.Address..Zip.Code.) == 5 & mergeDataAgg$TotalCount >= 100),]
mergeDataAgg$Percentage <- (mergeDataAgg$DefaultedCount/ mergeDataAgg$TotalCount) * 100
top10 <- head(mergeDataAgg[order(mergeDataAgg$Percentage, decreasing = T), c(1,4)], 10)
top10
```



```{r agg by amount}
#Agg by Amount
mergeDataAgg <- aggregate(mergeData[,c("TotalCount", "DefaultedCount")], by = list(mergeData$Total.Violation.Amount), "sum")
colnames(mergeDataAgg) <-c("Total.Violation.Amount", "TotalCount", "DefaultedCount")
mergeDataAgg <- mergeDataAgg[(mergeDataAgg$TotalCount >= 100),]
mergeDataAgg$Percentage <- (mergeDataAgg$DefaultedCount/ mergeDataAgg$TotalCount) * 100
top10 <- head(mergeDataAgg[order(mergeDataAgg$Percentage, decreasing = T), c(1,2, 4)], 10)
top10
```


# Clean Neighborhoods
```{r read shape file data}
library(rgdal)
nyc <- readOGR(dsn = "ZIP_CODE_040114", layer = "ZIP_CODE_040114")
```

```{r read clean violation data}
#Find out all violations related data from 
dataCleanLabel <- subset(data2015, Dirty.Label == 1 & Charge..1..Code.Description != "", select = c("Charge..1..Code.Description"))
library(tm)
dataCleanLabel <- as.vector(dataCleanLabel)
dataCleanCorpus <- VCorpus(VectorSource(dataCleanLabel))
dataCleanCorpus <- tm_map(dataCleanCorpus, removeWords, stopwords("english"))
dataCleanCorpus <- tm_map(dataCleanCorpus,  removePunctuation)
dataCleanCorpus <- tm_map(dataCleanCorpus, content_transformer(tolower))
dataCleanCorpus <- tm_map(dataCleanCorpus, stripWhitespace)

library(ngram)
library(RWeka)
UnigramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 1, max = 1))
dtm <- DocumentTermMatrix(dataCleanCorpus, control = list(tokenize = UnigramTokenizer, wordLengths = c(3,10)))

dtmFrame <- as.data.frame( t(as.matrix(  dtm )) ) 
dtmFrame$Term <- rownames(dtmFrame)
colnames(dtmFrame) <- c("count", "Term")

top30 <- head(dtmFrame[order(dtmFrame$count, decreasing = T), c(1, 2)], 30)
top30
#We uused unigram to verify words selected for selecting all the violations related to cleanliness

dataClean <- subset(data2015, Hearing.Result == "IN VIOLATION")
words <- c("clean", "receptacle", "dirty", "rubbish", "dirt", "waste", "garbage", "litter", "disposal", "dumping", "mosquito", "smoke")
dataClean <- dataClean[ifelse (rowSums(sapply(words, grepl, data2015$Charge..1..Code.Description, ignore.case = TRUE)) > 0, TRUE, FALSE),]
library(stringr)
dataClean$Charge..1..Code.Description <- str_trim(dataClean$Charge..1..Code.Description)

#Group by violation zip code and take count
dataCleanAgg <- do.call(data.frame, aggregate(cbind(count = Violation.Location..Zip.Code.) ~ Violation.Location..Zip.Code., data = dataClean, FUN = function(x) c({NROW(x)})))
colnames(dataCleanAgg) <- c("ZIPCODE", "count")
library(ggplot2)
library(Hmisc)
dataCleanAgg <- subset(dataCleanAgg, nchar(as.character(dataCleanAgg$ZIPCODE)) == 5)
dataCleanAgg$countGrp <- cut2(dataCleanAgg$count, c(0, 1, 30, 300, 1000, 4000))
```

```{r join with shape files}
#Left join with shape file data Ignoring Blank, 0 zipcodes from violation data
nyc@data <- merge(x = nyc@data, y = dataCleanAgg, by = "ZIPCODE", all.x = TRUE)
#Replacing NAs with group 1 to include 0 violations.
nyc@data <- data.frame(apply(nyc@data, 2, function(x){replace(x, is.na(x), "   0")}))
```

```{r plot map}
#Plot map
library(rgeos)
nyc_f <- fortify(nyc)
nyc$id <- row.names(nyc)
library(dplyr)
nyc_f <- left_join(nyc_f, nyc@data)
map <- ggplot() +
geom_polygon(data = nyc_f, aes(x = long, y = lat, group = group, fill = countGrp), color = "black", size = 0.25) +
coord_equal() +
labs(fill = "Number of Cleanliness\nRelated Violations 2015") +
ggtitle("Clean Neighborhoods")
map
```



```{r all dates}
#Generate all dates for year 2015
dateSeq <- data.frame(seq(as.Date("2015/01/01"), as.Date("2015/12/31"), by = "day"))
colnames(dateSeq) <- c("Violation.Date")
dateSeq$Violation.Date <- as.Date(dateSeq$Violation.Date, format = "%Y-%m-%d")
```

```{r pick top two zipcodes}
head(dataCleanAgg[order(as.numeric(dataCleanAgg$count), decreasing = T), ], 2)
```

```{r timeseries for zipcode 10019}
dataCleanSub <- subset(dataClean, Violation.Location..Zip.Code. == 10019, select = c("Violation.Date"))
dataCleanSub$Violation.Date <- as.Date(as.character(dataCleanSub$Violation.Date), "%m/%d/%Y")
dataCleanSubCnt <- data.frame(table(dataCleanSub))
colnames(dataCleanSubCnt) <- c("Violation.Date", "count")
dataCleanSubCnt$Violation.Date <- as.Date(dataCleanSubCnt$Violation.Date, format = "%Y-%m-%d")
mergeDateViolations <- merge(x = dateSeq, y = dataCleanSubCnt, by = "Violation.Date", all.x = TRUE)
#Replace NA with 0
mergeDateViolations <- data.frame(apply(mergeDateViolations, 2, function(x){replace(x, is.na(x), 0)}))
```

```{r plot ts for 10019}
#Daily
plot(ts(mergeDateViolations$count, start = c(2015, 1), frequency = 365), ylab = "count", main = "Zip Code 10019 Daily timeseries")
#Monthly
mergeDateViolations$Month <- 
format(as.Date(mergeDateViolations$Violation.Date), "%m-%Y")
mergeDateViolations$count <- as.numeric(as.character(mergeDateViolations$count))

mergeMonthViolations <- aggregate(mergeDateViolations[,c("count")], by=list(mergeDateViolations$Month), "sum")
colnames(mergeMonthViolations) <- c("Month", "count")

plot(ts(mergeMonthViolations$count, start = c(2015, 1), frequency = 12), ylab = "count", main = "Zip Code 10019 Monthly timeseries")
```

```{r decomposition for zip code 10019}
library(forecast)
weeklyAndMonthly <- msts(mergeDateViolations$count, seasonal.periods = c(7, 30, 365.25))
seasonalPattern <- tbats(weeklyAndMonthly)
#Zipcode 10019
plot(seasonalPattern)
```



```{r timeseries for zipcode 10003}
dataCleanSub <- subset(dataClean, Violation.Location..Zip.Code. == 10003, select = c("Violation.Date"))
dataCleanSub$Violation.Date <- as.Date(as.character(dataCleanSub$Violation.Date), "%m/%d/%Y")
dataCleanSubCnt <- data.frame(table(dataCleanSub))
colnames(dataCleanSubCnt) <- c("Violation.Date", "count")
dataCleanSubCnt$Violation.Date <- as.Date(dataCleanSubCnt$Violation.Date, format = "%Y-%m-%d")
mergeDateViolations <- merge(x = dateSeq, y = dataCleanSubCnt, by = "Violation.Date", all.x = TRUE)
#Replace NA with 0
mergeDateViolations <- data.frame(apply(mergeDateViolations, 2, function(x){replace(x, is.na(x), 0)}))
```

```{r plot ts for 10003}
#Daily
plot(ts(mergeDateViolations$count, start = c(2015, 1), frequency = 365), ylab = "count", main = "Zip Code 10019 Daily timeseries")
#Monthly
mergeDateViolations$Month <- 
format(as.Date(mergeDateViolations$Violation.Date), "%m-%Y")
mergeDateViolations$count <- as.numeric(as.character(mergeDateViolations$count))

mergeMonthViolations <- aggregate(mergeDateViolations[,c("count")], by=list(mergeDateViolations$Month), "sum")
colnames(mergeMonthViolations) <- c("Month", "count")

plot(ts(mergeMonthViolations$count, start = c(2015, 1), frequency = 12), ylab = "count", main = "Zip Code 10019 Monthly timeseries")
```

```{r decomposition for zip code 10003}
library(forecast)
weeklyAndMonthly <- msts(mergeDateViolations$count, seasonal.periods = c(7, 30, 365.25))
seasonalPattern <- tbats(weeklyAndMonthly)
#Zipcode 10019
plot(seasonalPattern)
```